---
title: "Exponential Distribution VS CLT"
author: "LP"
date: "Sunday, June 21, 2015"
output: pdf_document
---
This is the first part of the project for the Statistical Inference course, of the Data Science Specialization offered by the Johns Hopkins University.

##Overview

In this project we will investigate the exponential distribution in R and compare it with the Central Limit Theorem. 

We will attempt to:
1. Show the sample mean and compare it to the theoretical mean of the distribution.
2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
3. Show that the distribution is approximately normal.


##Simulations

The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. 
Set lambda = 0.2 for all of the simulations. We will investigate the distribution of averages of 40 exponentials. 

Here, we will do a thousand simulations, gather the means and take a peak at the top.

```{r}
#set the known variables and set a seed
n <- 40
no.sim <- 1000
lambda <- 0.2

set.seed(777)

#Simulation
sim_exp <- replicate(no.sim, rexp(n, lambda))

#Calculate means
means_exp <- apply(sim_exp, 2, mean)

#Look at the means
head(means_exp)
```

##Sample Mean VS Theoretical Mean
Here, we calculate the sample mean and the theoretical mean, to show where the distribution is centered.

```{r}
#calculate
sample_mean <- mean(means_exp)
theor_mean <- 1/lambda

#preview
sample_mean
theor_mean
```

We can observe that the center of the distribution is very close to the theoretical center of the distribution.

##Sample Variance VS Theoretical Variance
Now, we can calculate the sample variance and compare it to the theoretical variance.

```{r}
#calculate the standard deviation
theor_sd <- ((1/lambda) * (1/sqrt(n)))
sample_sd <- sd(means_exp)
         
#calculate the variance
theor_var <- theor_sd ^ 2
sample_var <- var(means_exp)

#preview the variance
theor_var
sample_var
```

Again, we observe that the variance of our sample and the expected variance are quite close.

##Distribution

Using a histogram, we can see observe our distribution of the means of the 1000 simulations.
The distribution is indeed nearly normal, with the red ink showing the theoretical distribution.



```{r}
library(ggplot2)

all_means <- data.frame(means_exp)

plott <- ggplot(all_means, aes (x = means_exp)) + 
  geom_histogram (binwidth = lambda, fill="yellow3", color="black", aes (y = ..density..)) + 
  labs(title="Exponential Distribution", x="Means", y="Density") + 
  stat_function(fun = dnorm, args = list(mean=theor_mean, sd=theor_sd), color = "red", size = 1.0)

print(plott)
```
