---
title: "ToothGroowth Analysis"
author: "LP"
date: "Sunday, June 21, 2015"
output: pdf_document
---

This is the second part of the project for the Statistical Inference course, of the Data Science Specialization offered by the Johns Hopkins University.


##Overview

For this part we are going to analyze the ToothGrowth data in the R datasets package.

The ToothGrowth dataset documents the effect of vitamin C on tooth growth in guinea pigs. The response is the length of odontoblasts (teeth) in each of 10 guinea pigs at each of three dose levels of Vitamin C (0.5, 1, and 2 mg) with each of two delivery methods (orange juice or ascorbic acid).

Further Information on the ToothGrowth dataset can be found here: https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/ToothGrowth.html


##Load the ToothGrowth data and perform some basic exploratory data analyses

First, we have to load the libraries we need and the ToothGrowth dataset.
Then, we can view part of the dataset (a dataframe with 60 observations on 3 variables) and by using boxplots we can perform a basic exploratory data analyses.

```{r}
#Load needed libraries
library(ggplot2)
library(datasets)
data(ToothGrowth)

#View the data
str(ToothGrowth)

head(ToothGrowth)

#Basic Analyses using Boxplots
plot <- ggplot(ToothGrowth, aes(x=factor(dose), y=len, fill=factor(dose))) + 
  geom_boxplot(notch=F) + facet_grid(.~supp) +
  scale_x_discrete("Dosage (in mg)") +   
  scale_y_continuous("Tooth Length") +  
  ggtitle("Supplement and Dosage Effects on Tooth Growth")
```

On the plot below, we can observe the effects of Vitamin C on the Tooth Growth of Guinea pigs. The dosage levels are 0.5, 1 and 2, while the delivery methods of vitamin C are Orange juice and Absorbic acid.

```{r, echo=FALSE}
plot
```



##Provide a basic summary of the data

Here, we have a summary of the ToothGrowth dataset, as well as a matrix for easy comparison between the dosage and the supplement.

```{r}
summary(ToothGrowth)

ToothGrowth$dose <- as.factor(ToothGrowth$dose)
table(ToothGrowth$supp, ToothGrowth$dose)
```


##Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose

We will now perform T-tests to see if there is a significant difference in tooth growth either by supplement type or by dosage variation.

```{r}
#Run a general T-test based on the type of supplement
t.test(len~supp, data=ToothGrowth)
```
This result (p-value=0.06 > 0.05) suggests that there is not much difference on tooth growth that is associated with the delivery method of Vitamin C.
However, we should take a closer look:

```{r}
#Run T-tests based specifically on each of the levels of dosage (0.5, 1, 2)
t.test(len~supp, data = subset(ToothGrowth, dose == 0.5))
t.test(len~supp, data = subset(ToothGrowth, dose == 1.0))
t.test(len~supp, data = subset(ToothGrowth, dose == 2.0))

```

The results from the individual T-tests that are based on specific dosage levels, suggest that  when it comes to lower dosages (dose of 0.5 mg and 1.0 mg) there is a significant difference on the tooth growth of guinea pigs that depends on the delivery method.
But for the larger dose (2.0 mg) the difference seems insignificant.


##State your conclusions and the assumptions needed for your conclusions

Gereral Assumptions to perform the tests above:
1. The sample of guinea pigs used is repesentative of their population 
2. There was random assignment of dosage and supplement on the guinea pigs

General Conclusions of the Analyses:
1. When the dosages are lower (0.5 mg & 1.0 mg), orange juice leads to more tooth growth than ascorbic acid.
2. When the dosage is higher (2.0 mg), the tooth growth is not statistically different between the two supplement delivery methods.
3. If we let aside the delivery method of the supplement, dosage is the factor that plays a key role in the tooth growth of guinea pigs.

